\documentclass[12pt]{exam}
\usepackage[utf8]{inputenc}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{multicol}
\usepackage{upquote}
\usepackage{url}
\newcommand{\class}{miniHive}
\newcommand{\term}{}
\newcommand{\examnum}{Setting up your own VM}
\usepackage{fancyvrb}
\newcommand{\examdate}{}

\pagestyle{head}
\firstpageheader{}{}{}
\runningheader{\class}{\examnum\ - Page \thepage\ of \numpages}{}
\runningheadrule



\begin{document}

\noindent
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r @{\extracolsep{6pt}} l}
\textbf{\class} & \\
\textbf{\term} &&\\
\textbf{\examnum} &&\\
\textbf{\examdate} &&\\
\end{tabular*}\\
\rule[2ex]{\textwidth}{2pt}

\noprintanswers



\begin{questions}

\question Register with Cloudera and download  the Cloudera Quickstart VM at \url{https://www.cloudera.com/downloads/quickstart_vms/5-13.html}.
The VM for VMWare worked best for me. 


\question
Spin up the VM. First of all, you may want to change the keyboard layout. Next, you may want to set up a shared folder with your host OS, so that you can easily share data.

This is how I mounted my shared folder. You may have to adapt the paths.

\begin{Verbatim}[fontsize=\footnotesize]
mkdir shares
sudo mount -t vmhgfs .host:/ ~/shares/    
\end{Verbatim}

\question
Open a  terminal. You'll need a more modern Python version and some extra modules.
This is what worked for me:

\begin{Verbatim}[fontsize=\footnotesize]
# Download Python 3.6
wget https://www.python.org/ftp/python/3.6.5/Python-3.6.5.tar.xz

xz -d Python-3.6.5.tar.xz
tar -xvf Python-3.6.5.tar

cd Python-3.6.5
./configure --prefix=/usr/local

# Let's build (compile) the source, this can take a while
make
sudo make altinstall
cd ..
sudo rm -rf Python*

# Install PIP    
wget https://bootstrap.pypa.io/get-pip.py
sudo /usr/local/bin/python3.6 get-pip.py
rm get-pip.py 
    
# Install further modules that we will need
sudo /usr/local/bin/python3.6 -m pip install luigi
sudo /usr/local/bin/python3.6 -m pip install sqlparse
sudo /usr/local/bin/python3.6 -m pip install radb 
sudo /usr/local/bin/python3.6 -m pip install pytest
sudo /usr/local/bin/python3.6 -m pip install pytest-repeat
\end{Verbatim}


\end{questions}

\end{document}
